---
title: "Ames_glm_boost"
output: html_notebook
---
Libraries
```{r}
library(plyr)
library(caret)
library(dplyr)
library(data.table)
```
Read in the data
```{r}
#Read the train/test file
initial_train_dt <- fread(input = "C:/Users/Chris Baskerville/Documents/data_sets/ames_data/train.csv", 
                  sep = ",", 
                  nrows = -1,
                  header = T,
                  na.strings = c("NA","N/A","null"),
                  stringsAsFactors = F,
                  check.names = T,
                  strip.white = T,
                  blank.lines.skip = T,
                  data.table = T
) 
graded_data_dt <- fread(input = "C:/Users/Chris Baskerville/Documents/data_sets/ames_data/test.csv", 
                 sep = ",", 
                 nrows = -1,
                 header = T,
                 na.strings = c("NA","N/A","null"),
                 stringsAsFactors = F,
                 check.names = T,
                 strip.white = T,
                 blank.lines.skip = T,
                 data.table = T
) 
```


```{r}
low <- unname(quantile(initial_train_dt$SalePrice, 0.25))
high <- unname(quantile(initial_train_dt$SalePrice, 0.75))
```
Compbine the data sets for joint manipulation

```{r}
## Create one data set for feature engineering. 
initial_train_dt[, dataPartition := "train"]
graded_data_dt[, SalePrice := as.integer(NA)] #Creates an SalePrice column in graded set containing NA values.
graded_data_dt[, dataPartition := "graded"]
comb_dt <- rbindlist(list(initial_train_dt, graded_data_dt), use.names = F, fill = F)
```

Fix typos
```{r}
comb_dt[YearRemodAdd > YrSold, YearRemodAdd := YrSold] ## Fix typo
comb_dt[GarageYrBlt == 2207, GarageYrBlt := 2007] ## Fix typo
comb_dt[MSSubClass  == 150, MSSubClass := 160] ## 150 not in training set
comb_dt[Exterior1st  == "Wd Sdng", Exterior1st := "WdSdng"] ## Fix spaces
comb_dt[Exterior2nd  == "Wd Sdng", Exterior2nd := "WdSdng"] ## Fix spaces
comb_dt[Exterior2nd  == "Brk Cmn", Exterior2nd := "BrkComm"] ## Fix typo
comb_dt[Exterior2nd  == "Wd Shng", Exterior2nd := "WdShing"] ## Fix typo
comb_dt[RoofMatl  == "Tar&Grv", RoofMatl := "TarGrv"] ## Fix '&'
comb_dt[RoofMatl  == "WdShngl", RoofMatl := "WdShing"] ## See exterior
```
Change some column names
```{r}
setnames(comb_dt, c("X1stFlrSF","X2ndFlrSF","X3SsnPorch"), c("FirstFlrSF","SecondFlrSF","ThreeSsnPorch"))
```


```{r}
variablesFactor <- colnames(comb_dt)[which(as.vector(comb_dt[,sapply(comb_dt, class)]) == "character")]
variablesFactor <- setdiff(variablesFactor, "dataPartition") 
variablesFactor <- c(variablesFactor,
                     ## variables with data type integer which are factors
                     "MSSubClass",     ## Identifies the type of dwelling involved in the sale
                     "OverallQual",    ## Rates the overall material and finish of the house
                     "OverallCond"     ## Rates the overall condition of the house
)
```

```{r}
variablesSquareFootage <- c(
  "LotFrontage", 		## Linear feet of street connected to property 
  "LotArea",    		## Lot size in square feet
  "MasVnrArea",  		## Masonry veneer area in square feet
  "BsmtFinSF1",		  ## Type 1 finished square feet	
  "BsmtFinSF2",		  ## Type 2 finished square feet
  "BsmtUnfSF",		  ## Unfinished square feet of basement area
  "TotalBsmtSF", 		## Total square feet of basement area
  "FirstFlrSF",		  ## First Floor square feet
  "SecondFlrSF",	  ## Second floor square feet
  "LowQualFinSF", 	## Low quality finished square feet (all floors)
  "GrLivArea", 		  ## Above grade (ground) living area square feet
  "GarageArea",     ## Size of garage in square feet
  "WoodDeckSF",     ## Wood deck area in square feet
  "OpenPorchSF",    ## Open porch area in square feet  
  "EnclosedPorch",  ## Enclosed porch area in square feet 
  "ThreeSsnPorch",  ## Three season porch area in square feet 
  "ScreenPorch",    ## Screen porch area in square feet
  "PoolArea" 		    ## Pool area in square feet
)
```
Count variable. Direct copy from Bart.
```{r}
variablesCounts <- c(
  "BsmtFullBath",		## Basement full bathrooms
  "BsmtHalfBath",		## Basement half bathrooms
  "FullBath",			  ## Full bathrooms above grade
  "HalfBath",			  ## Half baths above grade
  "BedroomAbvGr",		## Bedrooms above grade (does NOT include basement bedrooms)
  "KitchenAbvGr",		## Kitchens above grade
  "TotRmsAbvGrd",		## Total rooms above grade (does not include bathrooms)
  "Fireplaces",		  ## Number of fireplaces
  "GarageCars"     	## Size of garage in car capacity
)
```

```{r}
variablesValues <- c(
  "MiscVal",        ## $ Value of miscellaneous feature
  "SalePrice"       ## $ Price paid
)
```


```{r}
changeColType <- variablesFactor
comb_dt[,(changeColType) := lapply(.SD, as.factor), .SDcols = changeColType]
## Set columns to numeric
changeColType <- c(variablesSquareFootage, variablesCounts, variablesValues)
comb_dt[,(changeColType) := lapply(.SD, as.numeric), .SDcols = changeColType]
```
A function to find Columns containing NAs
```{r}
find_nas <- function(df){
  colnames(df)[apply(df, 2, anyNA)]
}
```
Catagorize what to do with NA columns
```{r}
#na catagories
Nones <- c('Alley', 'BsmtCond', 'BsmtExposure', 'BsmtFinType1', 'BsmtFinType2', 'BsmtQual', 'Exterior2nd', 'Fence', 'FireplaceQu', 'Functional', 'GarageCond', 'GarageFinish', 'GarageQual', 'GarageType', 'KitchenQual', 'MasVnrType', 'MiscFeature', 'PoolQC')

zeros <- c('BsmtFinSF1', 'BsmtFinSF2', 'BsmtFullBath', 'BsmtHalfBath', 'BsmtUnfSF', 'GarageArea', 'GarageCars', 'MasVnrArea', 'TotalBsmtSF' )

special <- c('GarageYrBlt', 'SaleType', 'Utilities', 'Exterior1st', 'Electrical', 'Exterior1st', 'MSZoning')
```
Imputation
```{r}
#impute nas in most numeric columns as 0 (easy: no column type
#conversion.Just find na and replace with the number 0)
#comb_dt[is.na(x), x := 0]
comb_dt[is.na(BsmtFinSF1), BsmtFinSF1 := 0]
comb_dt[is.na(BsmtFinSF2), BsmtFinSF2 := 0]
comb_dt[is.na(BsmtFullBath), BsmtFullBath := 0]
comb_dt[is.na(BsmtHalfBath), BsmtHalfBath := 0]
comb_dt[is.na(BsmtUnfSF), BsmtUnfSF := 0]
comb_dt[is.na(GarageArea), GarageArea := 0]
comb_dt[is.na(GarageCars), GarageCars := 0]
comb_dt[is.na(MasVnrArea), MasVnrArea := 0]
comb_dt[is.na(TotalBsmtSF), TotalBsmtSF := 0]

#Nones
#comb_dt[is.na(x), x := "None"]
comb_dt[is.na(Alley), Alley := "None"]
comb_dt[is.na(BsmtCond), BsmtCond := "None"]
comb_dt[is.na(BsmtExposure), BsmtExposure := "None"]
comb_dt[is.na(BsmtFinType1), BsmtFinType1 := "None"]
comb_dt[is.na(BsmtFinType2), BsmtFinType2 := "None"]
comb_dt[is.na(BsmtQual), BsmtQual := "None"]
comb_dt[is.na(Exterior2nd), Exterior2nd := "None"]
comb_dt[is.na(Fence), Fence := "None"]
comb_dt[is.na(FireplaceQu), FireplaceQu := "None"]
comb_dt[is.na(Functional), Functional := "None"]
comb_dt[is.na(GarageCond), GarageCond := "None"]
comb_dt[is.na(GarageFinish), GarageFinish := "None"]
comb_dt[is.na(GarageQual), GarageQual := "None"]
comb_dt[is.na(GarageType), GarageType := "None"]
comb_dt[is.na(KitchenQual), KitchenQual := "None"]
comb_dt[is.na(MasVnrType), MasVnrType := "None"]
comb_dt[is.na(MiscFeature), MiscFeature := "None"]
comb_dt[is.na(PoolQC), PoolQC := "None"]
  


  #Special: These are tough. Example: You can't impute a good year for
  #'GarageYrBlt' when there isn't one and 0 doesn't make sense in the
  #broader spectrum of things. Here, I'm going to use the year 1979 which
  #is the median of GarageYrBlt':
  
comb_dt[is.na(GarageYrBlt), GarageYrBlt := 1979]
  
  #'SaleType' Use the most common, which is "WD"
  
comb_dt[is.na(SaleType), SaleType := "WD" ]
  
  #"MSZoning" use the most common "RL"
comb_dt[is.na(MSZoning), MSZoning := "RL"]
  
  
  #Median value of LotFrontage is 68. There's reason to believe that
  #frontage isn't 0.
comb_dt[is.na(LotFrontage), LotFrontage := 68]
  
  #"Utilities" I'm going to drop this column, because they're all the same
  #except for the NAs and 1 without sewer and water.
comb_dt = subset(comb_dt, select = -Utilities)
  
  #"Exterior1st" Most common is 'VinylSd'
comb_dt[is.na(Exterior1st), Exterior1st := "VinylSd"]
  
  #"Electrical" Most common is 'SBrkr'
comb_dt[is.na(Electrical), Electrical := 'SBrkr']

find_nas(comb_dt)
#Should only be SalePrice for the Graded set.
```
reorder the factored data
```{r}
## OverallQual, rates the overall material and finish of the house
comb_dt[,OverallQual := ordered(OverallQual, levels = c(1:10))]
## OverallCond, rates the overall condition of the house
comb_dt[,OverallCond := ordered(OverallCond, levels = c(1:10))]
## KitchenQual, kitchen quality
comb_dt[,KitchenQual := ordered(KitchenQual, levels = c("None", "Po","Fa","TA","Gd","Ex"))]
## GarageFinish (contains NA's)
comb_dt[,GarageFinish := ordered(GarageFinish, levels = c("None","Unf","RFn","Fin"))]
## GarageQual
comb_dt[,GarageQual := ordered(GarageQual, levels = c("None","Po","Fa","TA","Gd","Ex"))]
## GarageCond
comb_dt[,GarageCond := ordered(GarageCond, levels = c("None","Po","Fa","TA","Gd","Ex"))]
## ExterQual, evaluates the quality of the material on the exterior  
comb_dt[,ExterQual := ordered(ExterQual, levels = c("Po","Fa","TA","Gd","Ex"))]
## ExterCond, evaluates the present condition of the material on the exterior
comb_dt[,ExterCond := ordered(ExterCond, levels = c("Po","Fa","TA","Gd","Ex"))]
## BsmtQual (contains NA's), evaluates the height of the basement
comb_dt[,BsmtQual := ordered(BsmtQual, levels = c("None","Po","Fa","TA","Gd","Ex"))]
## BsmtCond (contains NA's), evaluates the general condition of the basement
comb_dt[,BsmtCond := ordered(BsmtCond, levels = c("None","Po","Fa","TA","Gd","Ex"))]
## BsmtExposure (contains NA's), refers to walkout or garden level walls
comb_dt[,BsmtExposure := ordered(BsmtExposure, levels = c("None","No","Mn","Av","Gd"))]
## BsmtFinType1 (contains NA's), rating of basement finished area
comb_dt[,BsmtFinType1 := ordered(BsmtFinType1, levels = c("None","Unf","LwQ","Rec","BLQ","ALQ","GLQ"))]
## FireplaceQu (contains NA's), fireplace quality
comb_dt[,FireplaceQu := ordered(FireplaceQu, levels = c("None","Po","Fa","TA","Gd","Ex"))]
## Electrical
comb_dt[,Electrical := ordered(Electrical, levels = c("FuseP","Mix","FuseF","FuseA","SBrkr"))]
## Fence
comb_dt[,Fence := ordered(Fence, levels = c("None","MnWw","MnPrv","GdWo","GdPrv"))]
## PoolQC
comb_dt[,PoolQC := ordered(PoolQC, levels = c("None","Fa","Gd","Ex"))]
```

```{r}
nzv <- nearZeroVar(comb_dt, saveMetrics = T)
nzv[nzv$nzv,][1:21,]
```

```{r}
#comb_dt$BedroomAbvGr <- as.factor(comb_dt$BedroomAbvGr) #No Improvement here
#comb_dt$FullBath <- as.factor(comb_dt$FullBath) #No Improvement here
```


#```{r}
dim(comb_dt)
nzv <- nearZeroVar(comb_dt)
comb_dt[, (nzv) := NULL] 
dim(comb_dt)
#```

```{r}
dmy <- dummyVars(" ~ .", data = comb_dt)
comb_dt_dum <- data.table(predict(dmy, newdata = comb_dt))
```
Split the data into the original Training and Graded Test sets
```{r}
set.seed(42)
train_dt <- comb_dt[dataPartition == "train"]
train_dt[,dataPartition := NULL]
graded_dt <- comb_dt[dataPartition == "graded"]
graded_dt[,SalePrice := NULL]
graded_dt[,dataPartition := NULL]
find_nas(graded_dt)
```

Remove an outlier
```{r}
train_dt <- train_dt[GrLivArea < 4000]
```
Train / Test split on the training set
```{r}
set.seed(1)
trainIndex <- createDataPartition(train_dt$SalePrice, p = .8, 
                                  list = FALSE, 
                                  times = 1)
Ames_Train <- train_dt[ trainIndex,]
Ames_Test  <- train_dt[-trainIndex,]
```
Modeling using bayes lasso. This package has internal tuning parameters that take forever.
```{r}
library(doParallel)
library(monomvn)
ctrl <- trainControl(method = 'cv', number = 10,
                     classProbs = F,
                     summaryFunction = defaultSummary, verboseIter = TRUE) 

grid <- expand.grid(mstop = 1300, #seq(100 , 2000, by = 100),
                    prune = c('no'))
                   
model <- train(SalePrice ~. , data = Ames_Train,
                 method = 'blassoAveraged',
                 preProc = c("center", "BoxCox" ,"scale", 'nzv'), # "YeoJohnson" 'expoTrans' , "pca"
                 metric = "RMSE",
                 #tuneGrid = grid, #Allows us to pick the grid of training parameters for the optimal model selected by RMSE. Default tuning parameters choose three levels in the grid.
                 trControl = ctrl) 
```
RMSE Evaluation of the model
```{r}
Ames_pred <- predict(model, Ames_Test)
postResample(pred = Ames_pred, obs = Ames_Test$SalePrice)
```

```{r}
pred_comp <- data.frame(Ames_Test$SalePrice)
pred_comp$Preds <- Ames_pred
colnames(pred_comp)[1] <- "Actual"
ggplot(pred_comp, aes(y = Actual, x = Preds)) + geom_point() + ylim(0, 600000) +  xlim(0, 600000) + geom_abline(slope = 1, intercept = 0, col = "red")
```


Model Information:
```{r}
model
ggplot(model)
summary(model)
#coef(glmboostTune)
```

```{r}
glmboostPred <- predict(model, Ames_Test)
rmse = sqrt(apply((Ames_Test$SalePrice - glmboostPred)^2,2,mean))
rmse
plot(varImp(model), top = 20)
plot(model$mstop, rmse, type = 'b', xlab = "mstop")
```

```{r}
gradedPreds <- predict(glmboostTune, graded_dt)
submission <- data.frame(graded_dt$Id)
colnames(submission)[1] <- "Id"
submission$SalePrice <- as.vector(gradedPreds)
write.csv(submission, file = "C:/Users/Chris Baskerville/Documents/R_sessions and files/R_Projects/kaggle_ames_data/submission_0406_glmboost.csv", row.names = F)

```